---
title: "Parallel Programming in R"
output: html_document
date: "2022-11-01"
author: Danial Riaz and Ma.Fernanda Ortega
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```
## Install necessary packages

```{r}
#install.packages("tictoc")
#install.packages("future.apply")
#install.packages("furrr")
#install.packages("tidyverse")
#install.packages("stopwords")
#install.packages("quanteda")
#install.packages("quanteda.textstats")
```

## Load necessary libraries

```{r}
library(tidyverse)
library(tictoc)
library(parallel)
library(future.apply)
library(furrr)  
library(stopwords)
library(quanteda)
library(quanteda.textstats)
```

## Assess your own computer speed

Our ability to go parallel hinges on the number of CPU cores available to us. The simplest way to obtain this information from R is with the detectCores() function:

```{r}
detectCores()
```
This will indicate the number of CPU cores you have available on your computer to utilize and therefore how 'fast' your system can operate. You can adjust your expectations accordingly. 

## Exercise 1: "Tokenization (Serial implementation)"
For this exercise we will use a dataset that contains the titles of 23,481 fake news in order to separate the text into smaller units called tokens and remove words commonly used in the English language, such as "the", "is" and "and". 

```{r}

plan(sequential)
data_ML<- read_csv("C:/Users/feror/OneDrive/Documents/df_final_text.csv")## Download the file and choose your own path
tic()
mycorpus<-tokens(data_ML$text, 
             remove_punct = TRUE, # this removes punctuation
             remove_numbers = TRUE, # this removes digits
             remove_symbol = TRUE)%>% 
  tokens_remove(pattern = stopwords("en", source = "marimo"))
toc()

print(mycorpus[3])
```
## Exercise 1: "Tokenization (Parallel implementation with Future)"


```{r}

plan(multisession)
data_ML<- read_csv('C:\\Users\\feror\\Downloads\\df_final (1).csv')## Download the file and choose your own path
tic()
mycorpus<-future({tokens(data_ML$text, 
             remove_punct = TRUE, # this removes punctuation
             remove_numbers = TRUE, # this removes digits
             remove_symbol = TRUE)%>% 
  tokens_remove(pattern = stopwords("en", source = "marimo"))}) 
toc()
x<-value(mycorpus)
print(x[3])
```
### Conclusions Exercise 1: Tokenization
This example shows that the time it takes to tokenise and eliminate the "stop words" is significantly faster when using parallel implementation with Future. It should also be noted that only minor adjustments were made to the code to move to parallel processing. Furthermore, we will analyze the CPU history in our laptops to see what is happening while using Future package.



## Exercise 2: "Iterate over multiple inputs with Purrr" 

For this example we use the "unvotes" package that provides data on the voting history of countries in the United Nations General Assembly. This package contains three datasets: un_votes, providing the history of each country’s votes, un_roll_calls, providing information on each roll call vote, and un_roll_call_issues, providing issue (topic) classifications of roll call votes.

The first step is to create a function that takes country identifiers as well as a year_min argument as inputs and that returns the share of agreement in voting between any two specified countries as numeric value, for a time period specified with year >= year_min.

Secondly, we used the unique codes of the countries to apply the function "map_dbl" and find out which three countries on average agreed the most with the US from a given year.


```{r}
plan(sequential)
tic()
votes_agreement_calculator <- function(year_min, country1 = "", country2 = ""){
  
# votes country1
vote_decision_country1 <- unvotes::un_votes %>%
  filter(country_code == country1) %>%
  mutate(decision_country1 = vote) %>%
  select(rcid, decision_country1)

# votes country2
vote_decision_country2 <- unvotes::un_votes %>%
  filter(country_code == country2) %>%
  mutate(decision_country2 = vote) %>%
  select(rcid, decision_country2)

# get the year when a resolution happened
year_vote <- unvotes::un_roll_calls %>% 
  select(date, rcid) %>% 
  mutate(year = lubridate::year(date))

# combine data frames
un_votes_df <- 
vote_decision_country1 %>%
  left_join(vote_decision_country2, by = "rcid") %>%
  left_join(year_vote, by = "rcid") %>%
  filter(year >= year_min, !is.na(decision_country1), !is.na(decision_country2))

# calculate level of agreement between two countries
un_votes_df$agreement <- un_votes_df$decision_country1 == un_votes_df$decision_country2
agreement_share <- prop.table(table(un_votes_df$agreement))[2]
return(agreement_share)
}

country_codes_vec <- unvotes::un_votes %>% 
  pull(country_code) %>% 
  unique() %>% 
  na.omit() %>%
  as.character()

agreement_scores <- map_dbl(country_codes_vec, ~ votes_agreement_calculator(year_min = 2000, country1 = "US", country2 = .x))
toc()
data.frame(ccode = country_codes_vec, agree_share = agreement_scores) %>% arrange(desc(agree_share)) %>% slice_head(n = 3)
```

## Exercise 2: "Iterate over multiple inputs with Furrr" 

```{r}
plan(multisession)
tic()
votes_agreement_calculator <- function(year_min, country1 = "", country2 = ""){
  
# votes country1
vote_decision_country1 <- unvotes::un_votes %>%
  filter(country_code == country1) %>%
  mutate(decision_country1 = vote) %>%
  select(rcid, decision_country1)

# votes country2
vote_decision_country2 <- unvotes::un_votes %>%
  filter(country_code == country2) %>%
  mutate(decision_country2 = vote) %>%
  select(rcid, decision_country2)

# get the year when a decision happened
year_vote <- unvotes::un_roll_calls %>% 
  select(date, rcid) %>% 
  mutate(year = lubridate::year(date))

# combine data frames
un_votes_df <- 
vote_decision_country1 %>%
  left_join(vote_decision_country2, by = "rcid") %>%
  left_join(year_vote, by = "rcid") %>%
  filter(year >= year_min, !is.na(decision_country1), !is.na(decision_country2))

# calculate level of agreement between two countries
un_votes_df$agreement <- un_votes_df$decision_country1 == un_votes_df$decision_country2
agreement_share <- prop.table(table(un_votes_df$agreement))[2]
return(agreement_share)
}

country_codes_vec <- unvotes::un_votes %>% 
  pull(country_code) %>% 
  unique() %>% 
  na.omit() %>%
  as.character()

agreement_scores <- future_map_dbl(country_codes_vec, ~ votes_agreement_calculator(year_min = 2000, country1 = "US", country2 = .x))
toc()
data.frame(ccode = country_codes_vec, agree_share = agreement_scores) %>% arrange(desc(agree_share)) %>% slice_head(n = 3)

```

### Conclusion Exercise 2: Iterate over multiple inputs with Purrr and Furrr 

With the above example we managed to decrease the execution time by using the Furrr package. However, it should be noted that the code runs only a bit faster than when we used Purrr. This can be explained by Amdahl's Law, which states that there are diminishing returns in parallelization, depending on the proportion of your code that can be executed in parallel.


## Exercise 3: "Bootstrapping coefficient values for hypothesis testing (Serial implementation)"

For the last exercise we will create a fake data set (fake_data) and specifying a bootstrapping function (bootstrp()). This function will draw a sample of 10,000 observations from the the data set (with replacement), fit a regression, and then extract the coefficient on the x variable. 
```{r}
plan(sequential)
## Set seed (for reproducibility)
set.seed(1234)
# Set sample size
n = 1e6
tic()
## Generate a large data frame of fake data for a regression
  fake_data = 
  tibble(x = rnorm(n), e = rnorm(n)) %>%
  mutate(y = 3 + 2*x + e)

## Function that draws a sample of 10,000 observations, runs a regression and
## extracts the coefficient value on the x variable (should be around 2).
bootstrp = 
  function(i) {
  ## Sample the data
  sample_data = sample_n(fake_data, size = 1e4, replace = TRUE)
  ## Run the regression on our sampled data and extract the extract the x
  ## coefficient.
  x_coef = lm(y ~ x, data = sample_data)$coef[2]
  ## Return value
  return(tibble(x_coef = x_coef))
  }

## 10,000-iteration simulation
sim_serial = lapply(1:1e4, bootstrp) %>% bind_rows()
toc(log = TRUE)
head(sim_serial)
```

##  Exercise 3: "Bootstrapping coefficient values for hypothesis testing (Parallel implementation)"
```{r}
## Set seed (for reproducibility)
plan(multisession)
set.seed(1234)
# Set sample size
n = 1e6
tic()
## Generate a large data frame of fake data for a regression
  fake_data = 
  tibble(x = rnorm(n), e = rnorm(n)) %>%
  mutate(y = 3 + 2*x + e)

## Function that draws a sample of 10,000 observations, runs a regression and
## extracts the coefficient value on the x variable (should be around 2).
bootstrp = 
  function(i) {
  ## Sample the data
  sample_data = sample_n(fake_data, size = 1e4, replace = TRUE)
  ## Run the regression on our sampled data and extract the extract the x
  ## coefficient.
  x_coef = lm(y ~ x, data = sample_data)$coef[2]
  ## Return value
  return(tibble(x_coef = x_coef))
  }

## 10,000-iteration simulation
sim_serial = future_lapply(1:1e4, bootstrp,future.seed=123L) %>% bind_rows()
toc(log = TRUE)
head(sim_serial)
```

### Conclusion: Exercise 3 Bootstrapping coefficient values for hypothesis testing

This example shows that the time it takes to do the 10,000-iteration simulation is approximately 4 times faster when using parallel implementation with the package Future.apply. This example can also be evaluated with the package Furrr and we would have similar results. Furthermore, we will analyze the CPU history in our laptops to see what is happening while using Future package.










